<span style='line-height:60px; font-size:60px'>Changing communicative needs <br>and the structure of the lexicon:<br>corpus-based and experimental approaches</span></font>

### 2020-04-01 cletalk2020/index.Rmd
extra slide
new plots

@@

@@

```{r examplespace, out.width="95%", fig.height=3, eval=F}
#expl

library(ggplot2)
ex = data.frame(x={set.seed(1);rnorm(1000)}, y={set.seed(2);rnorm(1000)},change=rnorm(1000,sd=2.4) );ex$change[1]=10;ex$change[2]=-10; ex$change[200:1000] = rnorm(801, sd=0.2)
rownames(ex)=1:1000
cowplot::plot_grid(
ggplot(ex, aes(x,y, color=change)) + 
  annotate("line", x=ex[c(446, 656, 963),"x"], y=ex[c(446, 656, 963),"y"], color="gray") +
  geom_point(size=1.4, alpha=0.7) +
  geom_point(aes(x=ex[446,"x"], y=ex[446,"y"]), color=rgb(1,0,0), size=2.2) +
  geom_point(aes(x=ex[656,"x"], y=ex[656,"y"]), color="#2B2BF0", size=1.5) +
  geom_point(aes(x=ex[963,"x"], y=ex[963,"y"]), color="#8B8BCF", size=1.5)+
  theme_minimal(base_size = 10) + 
    theme(legend.position = "right") +
    scale_color_gradient2(low="blue", mid="lightgray", high="red") +
  labs(x="", y="") + theme(legend.position = "none")
,

ggplot(ex, aes(x,y, color=change)) + 
  annotate("line", x=ex[c(446, 656, 963),"x"], y=ex[c(446, 656, 963),"y"], color="black", size=1)+
  geom_point(size=4.4, alpha=0.7) +
  geom_point(aes(x=ex[446,"x"], y=ex[446,"y"]), color=rgb(1,0,0), size=9) +
  geom_point(aes(x=ex[656,"x"], y=ex[656,"y"]), color="#2B2BF0", size=4.6) +
  geom_point(aes(x=ex[963,"x"], y=ex[963,"y"]), color="#8B8BCF", size=4.6)+
  annotate("text", x=ex[c(446, 656, 963),"x"]+0.13, y=ex[c(446, 656, 963),"y"], label=c("+10", "-8", "-2"), color="black", size=3) +
  theme_minimal(base_size = 10) + 
    theme(legend.position = "right") +
    scale_color_gradient2(low="blue", mid="lightgray", high="red", name= "change\npmw") +
  labs(x="", y="") + lims(x=c(-3.2,-2.2), y=c(-0.5,0.4))
)

@@

- Approach: train word embeddings on period subcorpora (sliding window), scan for denser-than-expected subspaces - word clusters at a proximity where in 95% cases there would be no more than 1 word
--

@@

```{r, fig.width=12,fig.height=6 , out.height=600, out.width=1200,eval=T}
slides_beedemo
```
.small[Ordinal regression; McFadden's pseudo-R2; kappa = baseline-adjusted classification accuracy, cross-validated mean]

---
class: noborder
```{r, fig.width=12,fig.height=6 , out.height=600, out.width=1200,eval=T}
slides_beesplot
```
.small[Ordinal regression; McFadden's pseudo-R2; kappa = baseline-adjusted classification accuracy, cross-validated mean]

>>>>

<img src="img/tcm_anim.mov.gif" width="80%" />

@@

```{r examplespace, out.width="95%", fig.height=3, eval=T}
#expl

ex = data.frame(x={set.seed(1);rnorm(1000)}, y={set.seed(2);rnorm(1000)},change=rnorm(1000,sd=2.4) );ex$change[1]=10;ex$change[2]=-10; ex$change[200:1000] = rnorm(801, sd=0.2)
rownames(ex)=1:1000

e1 = ggplot(ex, aes(x,y, color=change)) + 
  annotate("line", x=ex[c(446, 656, 963),"x"], y=ex[c(446, 656, 963),"y"], color="gray") +
  geom_point(size=1.4, alpha=0.7) +
  geom_point(aes(x=ex[446,"x"], y=ex[446,"y"]), color=rgb(1,0,0), size=2.2) +
  geom_point(aes(x=ex[656,"x"], y=ex[656,"y"]), color="#2B2BF0", size=1.5) +
  geom_point(aes(x=ex[963,"x"], y=ex[963,"y"]), color="#8B8BCF", size=1.5)+
  theme_minimal(base_size = 10) + 
    theme(legend.position = "right") +
    scale_color_gradient2(low="blue", mid="lightgray", high="red") +
  labs(x="", y="") + theme(legend.position = "none")

e2 = ggplot(ex, aes(x,y, color=change)) + 
  annotate("line", x=ex[c(446, 656, 963),"x"], y=ex[c(446, 656, 963),"y"], color="black", size=1)+
  geom_point(size=4.4, alpha=0.7) +
  geom_point(aes(x=ex[446,"x"], y=ex[446,"y"]), color=rgb(1,0,0), size=9) +
  geom_point(aes(x=ex[656,"x"], y=ex[656,"y"]), color="#2B2BF0", size=4.6) +
  geom_point(aes(x=ex[963,"x"], y=ex[963,"y"]), color="#8B8BCF", size=4.6)+
  annotate("text", x=ex[c(446, 656, 963),"x"]+0.13, y=ex[c(446, 656, 963),"y"], label=c("+10", "-8", "-2"), color="black", size=3) +
  theme_minimal(base_size = 10) + 
    theme(legend.position = "right") +
    scale_color_gradient2(low="blue", mid="lightgray", high="red", name= "change\npmw") +
  labs(x="", y="") + lims(x=c(-3.2,-2.2), y=c(-0.5,0.4))

e1 + plot_spacer()

```{r examplespace2, out.width="95%", fig.height=3, eval=T}
e1 + e2
```

@@

@@

- Approach: train word embeddings on period subcorpora (sliding window), scan for denser-than-expected subspaces 
- word clusters at a proximity where in 95% cases there would be no more than 1 word

```{r, out.width="95%", fig.height=2, eval=T}
x={rnorm(10000,0.72,0.08) %>% subset(.<1)}
q = sort(x)[0.95*length(x)]

ggplot(data.frame(x),aes(y=x, x=1)) + 
  geom_violin(color="lightgray", fill="lightgray") +
  geom_boxplot(fill="lightgray", color="darkgray", outlier.color = "darkgray", width=0.6,outlier.size = 0.2) +
  geom_hline(yintercept = q, size=2)+
  ylim(0.25,1) + ylab("Cosine similarity of nearest neighbour") + 
 # labs(subtitle = "Nearest neighbor in the semantic space; entire lexicon") +
  annotate("text", x=1.5, y=0.83, label="95% of the words", hjust=1) +
  annotate("text", x=1.5, y=q+0.01, label="5% of the words", hjust=0) +
  annotate("text", x=0.5, y=0.83, label=paste0("Colexification threshold = ", round(q,2),"  ->"), hjust=1, fontface="bold") +
  theme_minimal() +
  theme(legend.position = "none", 
               panel.border = element_rect(color="darkgray",fill=NA), 
               axis.ticks = element_line(color="darkgray"), 
               strip.background = element_blank(),
               axis.text.x = element_text()
               #strip.text = element_blank(),
         )  +
         theme(#panel.grid.major.x = element_blank(),
               panel.grid.minor.y = element_blank(),
                panel.grid.minor.x = element_blank(),
               panel.grid.major.y = element_blank(),
               axis.text.y=element_blank(),
               axis.title.y=element_blank(),
               axis.ticks.y =element_blank()
         )  +
         theme(
            panel.border = element_rect(color="darkgray",fill=NA), 
            axis.ticks = element_line(color="darkgray"), 
            strip.background = element_blank(),
            strip.text = element_text(hjust=0, size = 13)
            #strip.text = element_blank(),
         ) +coord_flip()

@@

```{r, fig.width=12,fig.height=6 , out.height=600, out.width=1200,eval=T}
slides_beedemo
```
.small[Ordinal regression; McFadden's R^2; kappa = baseline-adjusted accuracy (10-fold CV). Improvement over reduced model in brackets.]

---
class: noborder
```{r, fig.width=12,fig.height=6 , out.height=600, out.width=1200,eval=T}
slides_beesplot
```
.small[Ordinal regression; McFadden's R^2; kappa = baseline-adjusted accuracy (10-fold CV). Improvement over reduced model in brackets.]
